### 2019-12-01 writing/manuscript.Rmd

---
title: "The effects of linkage disequilibrium on maintaining balanced polymorphisms in seasonal adaptation"
author: \emph{Thomas O'Leary, Csenge Petak, Alison Hall}
output: 
  pdf_document:
    number_sections: true
bibliography: refs.bib
header-includes:
   - \usepackage{graphicx}
   - \usepackage{float}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')

require(tidyverse)
require(cowplot)
source(here::here("functions.R"))
```

# Abstract

# Introduction
Something about an orchard in Pennsylvania [@Bergland2014]. As an example of how to cite.

# Methods

## Basic Model

The method for calculating fitness is based on the work of @Wittmann2017 who originally proposed _segregation lift_ as a general mechanism which may maintain balanced polymorphisms in a population under certain conditions.

The model first determines an individual's seasonal score ($z$) according to the equation:

$$
z = n_{homo} + d * n_{het}
$$
where $n_{homo}$ is the number loci homozygous for the allele particular to that certain season and $d$ is a dominance parameter (which ranges between 0 and 1) and $n_{het}$ is the number of heterozygous loci. A dominance parameter of greater than one ($d > 1$) would correspond to a traditional heterozygote advantage, and therefore is not of interest in this model.

The model we propose differs from the _segregation lift_ model in a few important ways. For one, the loci in our model are located on diploid linked chromosomes, so the homologous pairs can recombine during a simulated meiosis before mating. Expand on why this may be interesting or important… Linked genes not require a $d > 0.5$. Some genes that may be important for seasonal traits (_i.e._ heat shock proteins) are located close to each other on chromosomes and have a high amount of linkage disequilibrium…

Our model also does not allow for “selfing” where an individual can asexually reproduce with itself to create an offspring… Other stufff maybe too

| Default Parameters               |           |
|----------------------------------|-----------|
| Crossover probability            | 0.05      |
| Mutation probability             | $10^{-4}$ |
| Population size                  | 500       |
| Number of loci                   | 100       |
| Exponent of the fitness function | 1         |
| Dominance                        | 0.5       |
| Years                            | 100       |
| Generations per year             | 20        |
| Seasonal balance                 | 2         |

## Simulated seasonal selection

The general outline for the seasonal selection simulation:

1. A population of $N$ individuals is created with a single diploid chromosome containing $L$ loci, all of which contribute to the individual’s seasonal fitness. 
2. Seasonal score is calculated according to Eqn 1 and the fitness is determined
Pairs of parents – of size $N$ – are sampled stochasticly with a probability proportional to their seasonal fitness. 
3. Each parent in the pair undergoes recombination with a per locus recombination rate of $P_c$ 
4. A chromosome from each parent is selected at random to create the diploid offsprings 
5. The new generation replaces the old generation in the population.
6. The new individuals undergo mutation with a per locus mutation probability $P_m$ which was fixed at the relatively high rate of 1x10^-14 [@Haag-Liautard2007]. 
7. A new generation begins and the current season is determined by the current generation and seasonal balance parameter $S_b$. 
8. A new generation of parents is selected according to their seasonal fitness and the process continues to cycle for a number of years.
9. Loci specific allele frequencies are tracked throughout the simulated evolutionary run.

# Results

## Crossover rate

## Epistasis

This is how you reference Figure \ref{epi_density} in the text.

```{r, fig.height = 3, fig.width = 5, fig.cap = 'A density plot of the allele frequencies at the final generation for five replicates of each value of the epistatis parameter. \\label{epi_density}', message = FALSE, echo = FALSE}
setwd(here::here("results/epistasis"))

# load all files and filter only the final generation
all_tbl <- list.files(pattern = "c_0.05") %>% 
  map_df(~read_plus(.))

tbl_max_gen <- all_tbl %>%
  split(.$filename) %>%
  map(function(.data){filter(.data, genz == max(genz))}) %>%
  do.call("rbind", .)

# plot the density distribution
p1 <- ggplot(tbl_max_gen, aes(x = freqs, fill = filename)) +
        geom_density(alpha = 0.2) + 
        scale_fill_manual(values = c(rep("#e25dbb", 5), 
                                     rep("#6cdf4c", 5),
                                     rep("#258bd2", 5)),
                          breaks = unique(all_tbl$filename)[c(1,6,11)],
                          # legend name
                          name = "Epistasis",
                          # values tested
                          labels = c("0.5", "1", "2")) +
        ylab("Density") + 
        xlab("Allele Frequency") +
        theme_classic()

print(p1)

```

## Dominance

This is how you reference Figure \ref{dom_density} in the text.

```{r, fig.height = 3, fig.width = 5, fig.cap = 'A density plot of the allele frequencies at the final generation for five replicates of each value of dominance. \\label{dom_density}', message = FALSE, echo = FALSE}
setwd(here::here("results/dominance"))

# load all files and filter only the final generation
all_tbl <- list.files() %>% 
  map_df(~read_plus(.))

tbl_max_gen <- all_tbl %>%
  split(.$filename) %>%
  map(function(.data){filter(.data, genz == max(genz))}) %>%
  do.call("rbind", .)

# plot the density distribution
p2 <- ggplot(tbl_max_gen, aes(x = freqs, fill = filename)) +
        geom_density(alpha = 0.2) + 
        scale_fill_manual(values = c(rep("#e25dbb", 5), 
                                     rep("#6cdf4c", 5),
                                     rep("#258bd2", 5)),
                          breaks = unique(all_tbl$filename)[c(1,6,11)],
                          # legend name
                          name = "Dominance",
                          # values tested
                          labels = c("0.2", "0.5", "0.8")) +
        ylab("Density") + 
        xlab("Allele Frequency") +
        theme_classic()

print(p2)

```

## Generations per season

This is how you reference Figure \ref{gen_density} in the text.

```{r, fig.height = 3, fig.width = 5, fig.cap = 'A density plot of the allele frequencies at the final generation for five replicates of each number of generations per year. \\label{gen_density}', message = FALSE, echo = FALSE}
setwd(here::here("results/generations"))

# load all files and filter only the final generation
all_tbl <- list.files() %>% 
  map_df(~read_plus(.))

tbl_max_gen <- all_tbl %>%
  split(.$filename) %>%
  map(function(.data){filter(.data, genz == max(genz))}) %>%
  do.call("rbind", .)

# plot the density distribution
p3 <- ggplot(tbl_max_gen, aes(x = freqs, fill = filename)) +
        geom_density(alpha = 0.2) + 
        scale_fill_manual(values = c(rep("#e25dbb", 5), 
                                     rep("#6cdf4c", 5),
                                     rep("#258bd2", 5)),
                          breaks = unique(all_tbl$filename)[c(1,6,11)],
                          # legend name
                          name = "Generations\nper year",
                          # values tested
                          labels = c("10", "20", "50")) +
        ylab("Density") + 
        xlab("Allele Frequency") +
        theme_classic()

print(p3)
```

```{r, fig.height = 3, fig.width = 8 , fig.cap = 'A density plot of the allele frequencies at the final generation \\label{density_combined}', message = FALSE, echo = FALSE}

# this is how you could potentially mash a few figures together in the same plot
plot_grid(p1, p2, labels = "AUTO")
```

## Seasonal balance

## Population size

# Discussion

# References