### 2019-12-02 writing/manuscript.Rmd
Introduction and abstract

---
title: "The effects of linkage disequilibrium on maintaining balanced polymorphisms in seasonal adaptation"
author: \emph{Thomas O'Leary, Csenge Petak, Alison Hall}
output: 
  pdf_document:
    number_sections: true
bibliography: refs.bib
header-includes:
   - \usepackage{graphicx}
   - \usepackage{float}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')

require(tidyverse)
require(cowplot)
source(here::here("functions.R"))
```

# Abstract
Understanding the mechanisms of preserving genetic diversity is paramount to predicting the capacity of organisms to adapt to changing environmental conditions. In environments characterized by seasonal fluctuations, certain alleles may be beneficial under certain conditions and harmful in other environments. This fluctuation of additive benefits results in a strong directional selection that varies according to the season. @Bergland2014 experimentally observed that seasonal variation is able to preserve polymorphisms that maintain genetic diversity. @Wittman2017 modeled how allelic dominance changes according to the season in an organism to account for different seasonal life-history strategies. They present a term called “segregation lift” which describes instances of seasonal allele-fluctuations which serve to maintain genetic diversity over time. In this paper, we build upon the @Wittman2017 model, using the same parameters to study how genetic linkage and seasonal balance may affect segregation lift. The @Wittman2017 model found that in an unlinked genome the dominance parameter needed to be greater than 0.5 for segregation lift to occur. Our model suggests that in diploid organisms exhibiting genetic linkage, segregation lift occured when dominance=0.5, suggesting that multilocus polymorphisms are maintained if, at a certain locus, one allele over the other  is beneficial during one season.

# Introduction

Quantitative genetic theory suggests that plasticity occurs in response to selective pressure exerted by environmental heterogeneity [@Cooper2014].  Phenotypic plasticity is especially beneficial in variable environments because genetic variability allows for adaptation to changing conditions [@Gulisija2016]. Fluctuating selection can maintain polymorphisms in seasonal environments if strong selective pressure towards both homozygous genotypes varies seasonally [@Novak2017, @Bergland2014]. Understanding the role of seasonality on adaptation can allow more accurate predictions to be made about the capacity for organisms to adapt to climate change conditions [@Fournier-Level2016].

There has been some debate regarding whether balancing selection is capable of maintaining genetic polymorphisms in highly variable environments, and it is often thought that  this mechanism alone is not enough to maintain balanced polymorphism. (@Asthana2005, @Hedrick). However a study by @Bergland2014 observing Drosophila melanogaster in a temperate Pennsyvanian orchard revealed that strong seasonal selection can drive balancing selection favoring different life history traits over a short time span. These results suggest that seasonal variation contributes to the maintenance of two alleles to be present at the same loci long-term.  If there is no fitness cost associated with polymorphic variation and there is a benefit gained through heterozygosity, then it is suggested that that balancing selection could preserve genetic variation [@Asthana2005, @Wittman2017].

@Wittman2017 created a model simulating the seasonal selection observed by @Bergland2014 to clarify the role this allelic fluctuation plays in maintaining multilocus polymorphisms. This model introduces the term “segregation lift” which considers the importance of dominance in maintaining polymorphism across many loci in the genome and proposes that segregation lift plays an important role in preserving genetic variation.

Our model recreates and builds upon the work of @Wittman2017 to understand how linkage disequilibrium contributes to maintaining polymorphisms in a natural population. We study how rate of crossover, epistasis, allelic dominance, and seasonal balance contributes to segregation lift.

# Methods

## Basic Model

The method for calculating fitness is based on the work of @Wittmann2017 who originally proposed _segregation lift_ as a general mechanism which may maintain balanced polymorphisms in a population under certain conditions.

The model first determines an individual's seasonal score ($z$) according to the equation:

$$
z = n_{homo} + d * n_{het}
$$
where $n_{homo}$ is the number loci homozygous for the allele particular to that certain season and $d$ is a dominance parameter (which ranges between 0 and 1) and $n_{het}$ is the number of heterozygous loci. A dominance parameter of greater than one ($d > 1$) would correspond to a traditional heterozygote advantage, and therefore is not of interest in this model.

The model we propose differs from the _segregation lift_ model in a few important ways. For one, the loci in our model are located on diploid linked chromosomes, so the homologous pairs can recombine during a simulated meiosis before mating. Expand on why this may be interesting or important… Linked genes not require a $d > 0.5$. Some genes that may be important for seasonal traits (_i.e._ heat shock proteins) are located close to each other on chromosomes and have a high amount of linkage disequilibrium…

Our model also does not allow for “selfing” where an individual can asexually reproduce with itself to create an offspring… Other stufff maybe too

| Default Parameters               |           |
|----------------------------------|-----------|
| Crossover probability            | 0.05      |
| Mutation probability             | $10^{-4}$ |
| Population size                  | 500       |
| Number of loci                   | 100       |
| Exponent of the fitness function | 1         |
| Dominance                        | 0.5       |
| Years                            | 100       |
| Generations per year             | 20        |
| Seasonal balance                 | 2         |

## Simulated seasonal selection

The general outline for the seasonal selection simulation:

1. A population of $N$ individuals is created with a single diploid chromosome containing $L$ loci, all of which contribute to the individual’s seasonal fitness. 
2. Seasonal score is calculated according to Eqn 1 and the fitness is determined
Pairs of parents – of size $N$ – are sampled stochasticly with a probability proportional to their seasonal fitness. 
3. Each parent in the pair undergoes recombination with a per locus recombination rate of $P_c$ 
4. A chromosome from each parent is selected at random to create the diploid offsprings 
5. The new generation replaces the old generation in the population.
6. The new individuals undergo mutation with a per locus mutation probability $P_m$ which was fixed at the relatively high rate of 1x10^-14 [@Haag-Liautard2007]. 
7. A new generation begins and the current season is determined by the current generation and seasonal balance parameter $S_b$. 
8. A new generation of parents is selected according to their seasonal fitness and the process continues to cycle for a number of years.
9. Loci specific allele frequencies are tracked throughout the simulated evolutionary run.

# Results

## Crossover rate

## Epistasis

This is how you reference Figure \ref{epi_density} in the text.

```{r, fig.height = 3, fig.width = 5, fig.cap = 'A density plot of the allele frequencies at the final generation for five replicates of each value of the epistatis parameter. \\label{epi_density}', message = FALSE, echo = FALSE}
setwd(here::here("results/epistasis"))

# load all files and filter only the final generation
all_tbl <- list.files(pattern = "c_0.05") %>% 
  map_df(~read_plus(.))

tbl_max_gen <- all_tbl %>%
  split(.$filename) %>%
  map(function(.data){filter(.data, genz == max(genz))}) %>%
  do.call("rbind", .)

# plot the density distribution
p1 <- ggplot(tbl_max_gen, aes(x = freqs, fill = filename)) +
        geom_density(alpha = 0.2) + 
        scale_fill_manual(values = c(rep("#e25dbb", 5), 
                                     rep("#6cdf4c", 5),
                                     rep("#258bd2", 5)),
                          breaks = unique(all_tbl$filename)[c(1,6,11)],
                          # legend name
                          name = "Epistasis",
                          # values tested
                          labels = c("0.5", "1", "2")) +
        ylab("Density") + 
        xlab("Allele Frequency") +
        theme_classic()

print(p1)

```

## Dominance

This is how you reference Figure \ref{dom_density} in the text.

```{r, fig.height = 3, fig.width = 5, fig.cap = 'A density plot of the allele frequencies at the final generation for five replicates of each value of dominance. \\label{dom_density}', message = FALSE, echo = FALSE}
setwd(here::here("results/dominance"))

# load all files and filter only the final generation
all_tbl <- list.files() %>% 
  map_df(~read_plus(.))

tbl_max_gen <- all_tbl %>%
  split(.$filename) %>%
  map(function(.data){filter(.data, genz == max(genz))}) %>%
  do.call("rbind", .)

# plot the density distribution
p2 <- ggplot(tbl_max_gen, aes(x = freqs, fill = filename)) +
        geom_density(alpha = 0.2) + 
        scale_fill_manual(values = c(rep("#e25dbb", 5), 
                                     rep("#6cdf4c", 5),
                                     rep("#258bd2", 5)),
                          breaks = unique(all_tbl$filename)[c(1,6,11)],
                          # legend name
                          name = "Dominance",
                          # values tested
                          labels = c("0.2", "0.5", "0.8")) +
        ylab("Density") + 
        xlab("Allele Frequency") +
        theme_classic()

print(p2)

```

## Generations per season

This is how you reference Figure \ref{gen_density} in the text.

```{r, fig.height = 3, fig.width = 5, fig.cap = 'A density plot of the allele frequencies at the final generation for five replicates of each number of generations per year. \\label{gen_density}', message = FALSE, echo = FALSE}
setwd(here::here("results/generations"))

# load all files and filter only the final generation
all_tbl <- list.files() %>% 
  map_df(~read_plus(.))

tbl_max_gen <- all_tbl %>%
  split(.$filename) %>%
  map(function(.data){filter(.data, genz == max(genz))}) %>%
  do.call("rbind", .)

# plot the density distribution
p3 <- ggplot(tbl_max_gen, aes(x = freqs, fill = filename)) +
        geom_density(alpha = 0.2) + 
        scale_fill_manual(values = c(rep("#e25dbb", 5), 
                                     rep("#6cdf4c", 5),
                                     rep("#258bd2", 5)),
                          breaks = unique(all_tbl$filename)[c(1,6,11)],
                          # legend name
                          name = "Generations\nper year",
                          # values tested
                          labels = c("10", "20", "50")) +
        ylab("Density") + 
        xlab("Allele Frequency") +
        theme_classic()

print(p3)
```

```{r, fig.height = 3, fig.width = 8 , fig.cap = 'A density plot of the allele frequencies at the final generation \\label{density_combined}', message = FALSE, echo = FALSE}

# this is how you could potentially mash a few figures together in the same plot
plot_grid(p1, p2, labels = "AUTO")
```

## Seasonal balance

## Population size

# Discussion

# References